\section{Limitations and Scope}

BSC depends on the quality and coverage of the human-gold dimension set. If human
criteria omit a legitimate dimension, BSC cannot reward it. We therefore treat
RubricBench \texttt{test\_main} as the clean hard-gold benchmark and separate
it from proxy-gold training data. Proxy criteria elicited from teachers are used
for scale, but they should not be described as equivalent to human gold.

BSC is also a recall-style diagnostic over evaluation dimensions, not a full
measure of final judge correctness. Even when a method covers additional
human-gold dimensions, it may still use them poorly in a downstream decision,
or it can make a correct pairwise preference for the wrong reasons. For this
reason, BSC claims are paired with redundancy, hallucination, and downstream
utility gates;
the metric is intended to expose missing evaluative dimensions and test
attempted reductions, not to replace held-out judge-utility evaluation.
Concretely, an apparent coverage change is only metric-support unless the
RewardBench, RewardBench-2, and JudgeBench rows pass the paper-facing
API/model-scorer eligibility checks. If those rows are missing, stale, or
keyword-only smoke results, the correct manuscript sentence is metric-only BSC
evidence rather than downstream utility.

Semantic thresholds affect absolute BSC values. We report threshold sweeps and
confidence intervals, and the main claims should emphasize protocol-fixed
coverage changes rather than over-interpreting one threshold. A coverage change
should be described as dimension-level recovery only after the per-gold
dimension audit shows recovered and lost dimensions under query-aligned inputs,
fixed BGE embeddings, verifier valid-flag filtering, and the downstream utility gates.
Aggregate coverage alone cannot establish dimension-level recovery because it
can hide lost dimensions, duplicate paraphrases, or newly invalid criteria.
Embedding models can also introduce semantic matching errors. The repository
therefore provides an annotation-pack builder for matched and unmatched
dimension pairs, but the manuscript should treat it as C6-passing
human-audit evidence only after human labels are filled and audited. These safeguards make BSC
auditable rather than universal: they reduce threshold and embedding-model
arbitrariness, but they do not eliminate the need for downstream utility and
human-audit evidence.

The verifier is a necessary quality-control mechanism but may encode its own
biases. A strict verifier can wrongly reject valid domain-specific criteria,
while a relaxed verifier can admit generic or weak criteria. Our implementation
therefore uses fail-closed parsing, domain-aware policies, and separate
relaxed-verifier audits for HealthBench and BEIR/NQ. Paper claims about proxy
quality should be tied to verifier retention and validation reports.

Downstream utility depends on the criteria-conditioned scorer. Keyword overlap scorers are
only smoke-test tools; paper-facing downstream claims require a fixed
API/model scorer, consistent prompts, and held-out RewardBench, RewardBench-2,
and JudgeBench evaluation. Finally, this work studies whether LLMs cover human
evaluation dimensions, not whether human-authored criteria are complete or uniquely
correct for every open-ended task.

The contamination protocol is intentionally strict but not omniscient. Our
automated filters and audits operate at normalized query-level exact match, so
they are designed to catch direct training/evaluation leakage and known split
overlaps. They do not rule out every paraphrased, templated, or
semantically near-duplicate query across public corpora. We therefore report the
exact audit scope and require non-overlap holdouts for claims, while avoiding
stronger claims that no semantically related examples exist anywhere in the
training sources.
Conversely, a zero exact-overlap count is not a clean-isolation certificate if
the training-side artifacts are absent or stale. In that case the audit is
reported as \texttt{artifact\_status=blocked} and
\texttt{overlap\_status=not\_auditable}; the correct paper state is missing C0
evidence, not a clean holdout claim.
